Processing Code

### 2023-04-10 code/processing_code/processingcode.qmd
Part 4

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

@@

>>>>

### Load needed packages. make sure they are installed.
```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

@@

# Processing Section for Control Zip Code: 30907

## Processing 1: Getting Vaccination Totals by Control Zip Code 30907 in the Month of September

```{r}
#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Zip_30907_vax_Sept <-filter(Sept_17, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_vax_Sept)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_vax_Sept [Zip_30907_vax_Sept < 0] <- 0  
```

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Zip_30907_vax_Sept <- Zip_30907_vax_Sept  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian male residents 
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Zip_30907_vax_Sept<- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian female residents    
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

```{r}
#Vax total Asian Residents
Zip_30907_vax_Sept$total_asian_vax <- Zip_30907_vax_Sept$total_masian_vax + Zip_30907_vax_Sept$total_fasian_vax

#Vax total for African American Residents 
Zip_30907_vax_Sept$total_black_vax <- Zip_30907_vax_Sept$total_mblack_vax + Zip_30907_vax_Sept$total_fblack_vax

#Vax total for Other Race Residents
Zip_30907_vax_Sept$total_orace_vax <- Zip_30907_vax_Sept$total_morace_vax + Zip_30907_vax_Sept$total_forace_vax

#Vax total for White Residents
Zip_30907_vax_Sept$total_white_vax <- Zip_30907_vax_Sept$total_mwhite_vax + Zip_30907_vax_Sept$total_fwhite_vax  
```

```{r}
#summing all residents in the age range of 05-09
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```{r}
#Extract necessary variables only  
Zip_30907_vax_Sept_sums <- Zip_30907_vax_Sept[,c(1:4,185:204)]  
```

### Change the column names for 30907 Sept
```{r}
colnames(Zip_30907_vax_Sept_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Sept.rds")

saveRDS(Zip_30907_vax_Sept_sums, save_data_location4)
```

## Processing 2: Getting Vaccination Totals by Control Zip Code 30907 in the Month of October

```{r}
#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Zip_30907_vax_Oct <-filter(Oct_28, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_vax_Oct)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_vax_Oct [Zip_30907_vax_Oct < 0] <- 0  
```

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Zip_30907_vax_Oct <- Zip_30907_vax_Oct  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian male residents 
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Zip_30907_vax_Oct<- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian female residents    
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

```{r}
#Vax total Asian Residents
Zip_30907_vax_Oct$total_asian_vax <- Zip_30907_vax_Oct$total_masian_vax + Zip_30907_vax_Oct$total_fasian_vax

#Vax total for African American Residents 
Zip_30907_vax_Oct$total_black_vax <- Zip_30907_vax_Oct$total_mblack_vax + Zip_30907_vax_Oct$total_fblack_vax

#Vax total for Other Race Residents
Zip_30907_vax_Oct$total_orace_vax <- Zip_30907_vax_Oct$total_morace_vax + Zip_30907_vax_Oct$total_forace_vax

#Vax total for White Residents
Zip_30907_vax_Oct$total_white_vax <- Zip_30907_vax_Oct$total_mwhite_vax + Zip_30907_vax_Oct$total_fwhite_vax  
```

```{r}
#summing all residents in the age range of 05-09
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```{r}
#Extract necessary variables only  
Zip_30907_vax_Oct_sums <- Zip_30907_vax_Oct[,c(1:4,185:204)]  
```

### Change the column names for 30907 Oct
```{r}
colnames(Zip_30907_vax_Oct_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Oct.rds")

saveRDS(Zip_30907_vax_Oct_sums, save_data_location4)
```

## Processing 3: : Getting Vaccination Totals by Control Zip Code 30907 in the Month of November

```{r}
#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Zip_30907_vax_Nov <-filter(Nov_24, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_vax_Nov)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_vax_Nov [Zip_30907_vax_Nov < 0] <- 0  
```

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Zip_30907_vax_Nov <- Zip_30907_vax_Nov  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian male residents 
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Zip_30907_vax_Nov<- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian female residents    
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

```{r}
#Vax total Asian Residents
Zip_30907_vax_Nov$total_asian_vax <- Zip_30907_vax_Nov$total_masian_vax + Zip_30907_vax_Nov$total_fasian_vax

#Vax total for African American Residents 
Zip_30907_vax_Nov$total_black_vax <- Zip_30907_vax_Nov$total_mblack_vax + Zip_30907_vax_Nov$total_fblack_vax

#Vax total for Other Race Residents
Zip_30907_vax_Nov$total_orace_vax <- Zip_30907_vax_Nov$total_morace_vax + Zip_30907_vax_Nov$total_forace_vax

#Vax total for White Residents
Zip_30907_vax_Nov$total_white_vax <- Zip_30907_vax_Nov$total_mwhite_vax + Zip_30907_vax_Nov$total_fwhite_vax  
```

```{r}
#summing all residents in the age range of 05-09
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```{r}
#Extract necessary variables only  
Zip_30907_vax_Nov_sums <- Zip_30907_vax_Nov[,c(1:4,185:204)]  
```

### Change the column names for 30907 Nov
```{r}
colnames(Zip_30907_vax_Nov_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Nov.rds")

saveRDS(Zip_30907_vax_Nov_sums, save_data_location4)
```

## Processing 4: : Getting Vaccination Totals by Control Zip Code 30907 in the Month of December

```{r}
#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Zip_30907_vax_Dec <-filter(Dec_29, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_vax_Dec)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_vax_Dec [Zip_30907_vax_Dec < 0] <- 0  
```

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Zip_30907_vax_Dec <- Zip_30907_vax_Dec  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian male residents 
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Zip_30907_vax_Dec<- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```{r}
#total vax column for Asian female residents    
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

```{r}
#Vax total Asian Residents
Zip_30907_vax_Dec$total_asian_vax <- Zip_30907_vax_Dec$total_masian_vax + Zip_30907_vax_Dec$total_fasian_vax

#Vax total for African American Residents 
Zip_30907_vax_Dec$total_black_vax <- Zip_30907_vax_Dec$total_mblack_vax + Zip_30907_vax_Dec$total_fblack_vax

#Vax total for Other Race Residents
Zip_30907_vax_Dec$total_orace_vax <- Zip_30907_vax_Dec$total_morace_vax + Zip_30907_vax_Dec$total_forace_vax

#Vax total for White Residents
Zip_30907_vax_Dec$total_white_vax <- Zip_30907_vax_Dec$total_mwhite_vax + Zip_30907_vax_Dec$total_fwhite_vax  
```

```{r}
#summing all residents in the age range of 05-09
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
